Combining living and fossil taxa into phylogenies: the missing data issue

### 2015-01-12 Manuscript/TEM_draft.tex

%NC: Supp material - add the actual table and figures for each thing. Also BE CONSISTENT. Is it Section 1 or S1? I'd go with S1 (which stands for Supp Mat 1 not section 1 really.)

@@

We created matrices using all parameter combinations resulting in 125 ($5^3$) matrices. Note that one of these combinations has no missing data so is equivalent to the "complete" matrix, thus we have one effectively complete matrix in our 125 "missing-data" matrices. Because some parameter combinations introduce a lot of missing data (e.g. $M_L$=75\%, $M_F$=75\% and $M_C$=75\%), some matrices contained fossil taxa without any data at all. When this occurred we repeated the random deletion of characters until each taxon had at least 5\% data across the whole morphological part of the matrix.

@@

\begin{figure} 
\centering
    \includegraphics[width=1\textwidth]{Figures/In_main/AllMethods-RF+Tr-colour.pdf}
\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue). The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis. Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row). The graph shows the modal value (points), the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
\label{Fig_Results-permeth_perparam} %Differences within a subset of parameters and between methods
\end{figure}

\begin{figure} 
\centering
    \includegraphics[width=1\textwidth]{Figures/In_main/ML+Baycon-AllParam-RF+Tr-BW.pdf}
\caption{Effect of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey). The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on the x axis. Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row). The graph shows the modal value (points), the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.} 
\label{Fig_Results-global_perparam} %Differences between all the parameters and between two methods (ML vs BC)
\end{figure}

\subsection{Individual effects of missing data parameters}
As the amount of missing data increases across all three parameters, our ability to recover the "best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} Table 2, 3 and 4).
However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3} Table 2, 3 and 4).

\subsection{Combined effect of missing data parameters}
As expected, our ability to recover the "best" tree topology is worst when each parameter contains the maximum amount of missing data (i.e. $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%), and best when there is no missing data (i.e. $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}). Figure ~\ref{Fig_Results-paircomp_within} shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0; Fig.~\ref{Fig_Results-paircomp_within}).

\begin{figure} 
\centering
     %Hide this figure for 'fast' build
%    \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.pdf}
    \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
\caption{Effect of missing data on topological recovery using Bayesian consensus trees. The percentage of missing data (gradient colour from 0\% (white), 10\%, 25\%, 50\% and 75\% (black)) for each parameter ($M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$(lower line)) is shown on both x and y axis. Topological recovery is represented by the probability of the two different tree comparison metrics (Normalised Robinson-Foulds distance (A - right) and Normalised Triplets distance (B - left)) distribution overlap with the "best" tree using the the Bhattacharyya Coefficients (low probability of overlap - blue ; high probability of overlap - orange).}
\label{Fig_Results-paircomp_within}
\end{figure} %Pairwise BC for the Bayesian consensus trees for RF and TR. (parameters differences)

Using both Normalised Robinson-Foulds and Normalised Triplets distances from the Bayesian consensus trees, the parameter combination with no missing data (i.e. $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameter combinations (Fig.~\ref{Fig_Results-paircomp_within}). However, the Normalised Robinson-Foulds distance (median Bhattacharrya coefficient = 0.79 ; Fig.~\ref{Fig_Results-paircomp_within}-A) displays more dissimilarities than the Normalised Triplets distance (median Bhattacharrya coefficient = 0.81; Fig.~\ref{Fig_Results-paircomp_within}-B).Additionally, when using the Normalised Robinson-Foulds distance, once $M_{L}$ $\geq$  50\%, there is no additional affect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Fig.~\ref{Fig_Results-paircomp_within}-A). Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional affect of $M_{L}$ and $M_{F}$ (Fig.~\ref{Fig_Results-paircomp_within}-A).

For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.98, using Normalised Robinson-Foulds distance or Normalised Triplets distance respectively; see Table ~\ref{Tab_Results-Difference_methods}). However, these two methods perform worse than the Bayesian consensus trees sing Normalised Robinson-Foulds distance (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; see Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).

\begin{landscape}
\begin{table}[ht]
\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data".}
\centering
\begin{tabular}{lccccccc}
  \hline
 Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF^1$ & 0.00 & 0.00 & 0.10 & 0.20 & 0.32 & 1.00 \\ 
                                                                       & $Tr^2$ & 0.34 & 0.49 & 0.61 & 0.62 & 0.75 & 1.00 \\ 
    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
                                                                       & $Tr^2$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
                                                                       & $Tr^2$ & 0.21 & 0.67 & 0.73 & 0.72 & 0.77 & 0.84 \\ 
    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF^1$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
                                                                       & $Tr^2$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF^1$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
                                                                       & $Tr^2$ & 0.21 & 0.36 & 0.56 & 0.55 & 0.74 & 0.87 \\ 
    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF^1$ & 0.50 & 0.77 & \textbf{0.85} & \textbf{0.85} & 0.96 & 1.00 \\ 
                                                                       & $Tr^2$ & 0.91 & 0.96 & \textbf{0.98} & \textbf{0.97} & 0.99 & 1.00 \\ 
   \hline
\end{tabular}
   $^1RF$=Normalised Robison-Fould distance ; $^2Tr$=Normalised Triplets distance
\label{Tab_Results-Difference_methods}
\end{table}
\end{landscape}

@@

Our results show that the ability to recover the "best" tree topology in a Total Evidence framework decreases with increasing amount of missing data, regardless of how data were removed or the method of tree inference used. However, these factors affected topological recovery in different ways and to different extents. 
Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($M_{C}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}). Additionally, using Bayesian consensus trees recovered the "best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.~\ref{Fig_Results-global_perparam} and Fig.~\ref{Fig_Results-paircomp_within}). As shown in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.

@@

Additionally, the number of morphological characters determines the size the matrix. This can affect our ability to recover the "best" tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy. The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \citep{wagner2000}. The size of our data matrices were constrained by the performance of our protocol: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters. Therefore, part of the decrease of the Normalised Robinson-Fould distance and the Normalised Triplets distance in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}). However, although these matrices are an order of magnitude smaller than some published matrices \citep[e.g.][]{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices \citep[e.g.][]{kellymolecular2014, sallam2011craniodental}. Therefore, our simulations reflect realistic parameters. Homoplasy, on the other hand, is expected to increase with an increase of the morphological characters \citep{}. However, the use of probabilistic methods (i.e. Maximum Likelihood or Bayesian) and the M\textit{k} model \citep{lewisa2001} has been previously demonstrated to partially resolve this issue \citep{wrightbayesian2014}.

\subsection{Combined effect of missing data parameters}
As expected, when combining the missing data parameters, our ability to recover the "best" tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds distance and the Normalised Triplets distance are higher when all the missing data parameters have low missing data (i.e. $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have high missing data (i.e. $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Fig.~\ref{Fig_Results-global_perparam}). However, it is important to notice that the effect of each parameter is not additive. The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$); when both $M_{L}$ and $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Fig.~\ref{Fig_Results-paircomp_within}). This has important implications for building Total Evidence phylogenies (see below).

@@

Another practical implication of our results regards the tree inference methods and their ability to recover the "best" tree topology with missing data in a Total Evidence framework. Because the Bayesian consensus tree consistently recovered topologies closer to the "best" tree topology than any other methods, we advice to choose this tree topology as a fixed topology for any further step in tree inference analysis such as Tip-dating \citep{BEASTmaster}. However, although the posterior distribution tree distribution performed poorly in recovering the "best" tree topology, this is more an effect of a caveat in our methods (see discussion above). Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed since they can be useful for further comparative analysis.% \citep[e.g.][]{}. %Bayesian PCM cite

@@

Using Total Evidence matrices with missing data is not a problem for recovering the "best" tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters. When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $M_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa. Therefore we advise filling as many gaps in Total Evidence matrices as possible. Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters as possible and to code them for as many living taxa present in the matrix as possible. Additionally, the topology of the Bayesian consensus tree, regardless the amount of missing data, was always the closer to the "best" tree topology than the Maximum likelihood trees. Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed in a Total Evidence analysis, for example when using the Tip-Dating method. The results of this analysis are encouraging and showing that it is possible to combine both neontological and palaeontological data in the same analysis. Hopefully, using these approaches, will greatly improve our understanding of macroevolutionary processes through both time and space.

@@

\section{Acknowledgements}
Thanks to Gavin Thomas, Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson, Andrew Jackson, Gavin Thomas and April Wright for useful comments on our simulation protocol. Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster. Simulations used the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing and funded through grants from Science Foundation Ireland. This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).

@@

%---------------------------------------------
%
%       Supplementary, separated
%
%---------------------------------------------

%Ignore this for 'fast' pdf building
\section{Supplementary Material}

    \input{SupplementaryMaterial/SupplementaryMaterial}
    \label{SupplementaryMaterial}

>>>>

%NC: Supp material - add the actual table and figures for each thing. Also BE CONSISTENT. Is it Section 1 or S1? I'd go with S1 (which stands for Supp Mat 1 not section 1 really.). Normally you also refer to supplemental tables and figures as Table S1 etc. to distinguish between those in the main text. Check how Syst Biol does it.

@@

We created matrices using all parameter combinations resulting in 125 ($5^3$) matrices. Note that one of these combinations has no missing data so is equivalent to the "complete" matrix, thus we have one effectively complete matrix in our 125 "missing-data" matrices. Because some parameter combinations introduce a lot of missing data (e.g. $M_L$ = 75\%, $M_F$ = 75\% and $M_C$ = 75\%), some matrices contained fossil taxa without any data at all. When this occurred we repeated the random deletion of characters until each taxon had at least 5\% data across the whole morphological part of the matrix.

@@

\begin{figure} 
\centering
    \includegraphics[width=1\textwidth]{Figures/In_main/AllMethods-RF+Tr-colour.pdf}
\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black), Bayesian consensus trees (blue), Maximum Likelihood bootstrap trees (orange) and Bayesian posterior tree distributions (blue). The percentage of missing data for each parameter ($M_{L}$, $M_{F}$ and $M_{C}$) is shown on the x axis. Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row). The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.}
\label{Fig_Results-permeth_perparam} %Differences within a subset of parameters and between methods
\end{figure}

\begin{figure} 
\centering
    \includegraphics[width=1\textwidth]{Figures/In_main/ML+Baycon-AllParam-RF+Tr-BW.pdf}
\caption{The effects of increasing missing data on topological recovery using Maximum Likelihood trees (black) and Bayesian consensus trees (grey). The x axis shows the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line). Topological recovery was measured using two different tree comparison metrics: Normalised Robinson-Foulds distance (upper row) and Normalised Triplets distance (lower row). The graph shows the modal value (points), and the 50\% (thick solid lines) and 95\% (thin dashed lines) confidence intervals of the distributions of the tree comparison metric for each missing data parameter and tree inference method.} 
\label{Fig_Results-global_perparam} %Differences between all the parameters and between two methods (ML vs BC)
\end{figure}

\subsection{Individual effects of missing data parameters}
As the amount of missing data increases across all three parameters, our ability to recover the "best" tree topology decreases (Fig.~\ref{Fig_Results-permeth_perparam}).
The Normalised Robinson-Foulds distance is always lower for the Maximum Likelihood trees than for the Bayesian consensus trees (median Bhattacharrya Coefficient = 0.69, 0.48 and 0.66 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}: Tables S2-S4).
However, The Normalised Triplets distance is similar between the Maximum Likelihood trees and the Bayesian consensus trees for all the parameters ($M_{L}$, $M_{F}$ and $M_{C}$) (median Bhattacharrya Coefficient = 0.84, 0.75 and 0.80 for $M_{L}$, $M_{F}$ and $M_{C}$ respectively; Fig.~\ref{Fig_Results-permeth_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}: Tables S2-S4).

\subsection{Combined effect of missing data parameters}
As expected, our ability to recover the "best" tree topology is worst when each parameter contains the maximum amount of missing data (i.e. $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%), and best when there is no missing data (i.e. $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%; Fig.~\ref{Fig_Results-global_perparam}; \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}). Fig. ~\ref{Fig_Results-paircomp_within} shows the similarity of distributions of tree metrics in a triangular matrix with the values of each pairwise Bhattacharyya Coefficient coloured according to their values (orange when the distributions overlap completely, Bhattacharyya Coefficient = 1, and blue when they do not, Bhattacharyya Coefficient = 0; Fig.~\ref{Fig_Results-paircomp_within}).

\begin{figure} 
\centering
     %Hide this figure for 'fast' build
%    \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.pdf}
    \includegraphics[width=1\textwidth]{Figures/In_main/PairwiseComp-Baycon-RF+Tr-colour.png} %bitmap version - 300 dpi RGB
\caption{The effects of missing data on topological recovery using Bayesian consensus trees. Both axes show the percentage of missing data from 0\% (white) to 75\% (black) for the three parameters: $M_{L}$ (upper line), $M_{F}$ (middle line) and $M_{C}$ (lower line). Topological recovery is represented by the probability of (A) Normalised Robinson-Foulds distance and (B) Normalised Triplets distance distributions overlapping with the "best" tree distribution, calculated using the Bhattacharyya Coefficient. The Bhattacharyya Coefficient values indicating low probability of overlap are shown in blue, those indicating a high probability of overlap are shown in orange.}
\label{Fig_Results-paircomp_within}
\end{figure} %Pairwise BC for the Bayesian consensus trees for RF and TR. (parameters differences)

Using both Normalised Robinson-Foulds and Normalised Triplets distances from the Bayesian consensus trees, the parameter combination with no missing data (i.e. $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) is always the most dissimilar to all the other parameter combinations (Fig.~\ref{Fig_Results-paircomp_within}). However, the Normalised Robinson-Foulds distance (median Bhattacharrya coefficient = 0.79; Fig.~\ref{Fig_Results-paircomp_within}A) displays more dissimilarities than the Normalised Triplets distance (median Bhattacharrya coefficient = 0.81; Fig.~\ref{Fig_Results-paircomp_within}B). Additionally, when using the Normalised Robinson-Foulds distance, once $M_{L}$ $\geq$  50\%, there is no additional affect of $M_{F}$ and $M_{C}$, regardless of the amount of missing data in these parameters (Fig.~\ref{Fig_Results-paircomp_within}A). Likewise, once $M_{C}$ $\geq$ 50\%, there is no additional affect of $M_{L}$ and $M_{F}$ (Fig.~\ref{Fig_Results-paircomp_within}A).

For all combinations of missing data parameters and tree comparison metrics, the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distributions perform very similarly (median Bhattacharrya Coefficient = 0.85 and 0.98, using Normalised Robinson-Foulds distance or Normalised Triplets distance respectively; Table ~\ref{Tab_Results-Difference_methods}). However, these two methods perform worse than the Bayesian consensus trees using Normalised Robinson-Foulds distance (median Bhattacharrya Coefficient = 0 and 0.01, for the Maximum Likelihood bootstrap trees and the Bayesian posterior tree distribution respectively; Table ~\ref{Tab_Results-Difference_methods}; Fig.~\ref{Fig_Results-permeth_perparam} and \hyperref[SupplementaryMaterial]{Supplementary Material Section 3}).

% NC: In the caption you need to explain why some of these values are in bold...
\begin{landscape}
\begin{table}[ht]
\caption{Bhattacharyya Coefficients of the pairwise method comparisons, each of which corresponds to the normalised distance between the "best" tree and the "missing data" using either the Normalised Robinson-Foulds distance (RF) or the Normalised Triplets distance (Tr).}
\centering
\begin{tabular}{lccccccc}
  \hline
 Comparison &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
    Maximum Likelihood \textit{vs.} Bayesian consensus                 & $RF$ & 0.00 & 0.00 & 0.10 & 0.20 & 0.32 & 1.00 \\ 
                                                                       & $Tr$ & 0.34 & 0.49 & 0.61 & 0.62 & 0.75 & 1.00 \\ 
    Maximum Likelihood \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
                                                                       & $Tr$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
    Maximum Likelihood \textit{vs.} Bayesian posterior trees           & $RF$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
                                                                       & $Tr$ & 0.21 & 0.67 & 0.73 & 0.72 & 0.77 & 0.84 \\ 
    Bayesian consensus \textit{vs.} Maximum Likelihood bootstraps      & $RF$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
                                                                       & $Tr$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
    Bayesian consensus \textit{vs.} Bayesian posterior trees           & $RF$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
                                                                       & $Tr$ & 0.21 & 0.36 & 0.56 & 0.55 & 0.74 & 0.87 \\ 
    Bayesian posterior tree \textit{vs.} Maximum Likelihood bootstraps & $RF$ & 0.50 & 0.77 & \textbf{0.85} & \textbf{0.85} & 0.96 & 1.00 \\ 
                                                                       & $Tr$ & 0.91 & 0.96 & \textbf{0.98} & \textbf{0.97} & 0.99 & 1.00 \\ 
   \hline
\end{tabular}
\label{Tab_Results-Difference_methods}
\end{table}
\end{landscape}

@@

Our results show that the ability to recover the "best" tree topology in a Total Evidence framework decreases as the amount of missing data increases, regardless of how data were removed or the method of tree inference used. However, these factors affected topological recovery in different ways and to different extents. 
Decreasing the amount of living taxa with morphological data ($M_{L}$) and the overall number of morphological characters in the matrix ($M_{C}$) had worst effects on topological recovery (Fig.~\ref{Fig_Results-paircomp_within}). Additionally, using Bayesian consensus trees recovered the "best" tree topology more consistently than using Maximum Likelihood trees or Bayesian posterior tree distributions (Fig.~\ref{Fig_Results-global_perparam}, Fig.~\ref{Fig_Results-paircomp_within}, Table 1). As seen in previous studies, our results show that the amount of missing data is not a problem \textit{per se} for Total Evidence methods, as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters \citep[e.g.][]{kearneyfragmentary2002,wiensmissing2003,rouresite-specific2011,pattinsonphylogeny2014}.

@@

Additionally, the number of morphological characters determines the size the matrix. This can affect our ability to recover the "best" tree topology through: (1) the incongruence of phylogenetic signal among morphological and molecular data; and/or (2) homoplasy. The incongruence of phylogenetic signal between morphological and molecular data has previously been demonstrated to be more important in small morphological matrices \citep{wagner2000}. The size of our data matrices were constrained by the performance of our protocol: to reduce the computational time of our analysis to a reasonable level (150 CPU years), we ran our simulations on modestly-sized matrices of 1000 molecular characters and 100 morphological characters. Therefore, part of the decrease of the Normalised Robinson-Fould distance and the Normalised Triplets distance in our simulations could be due to conflicting phylogenetic signal among morphological and molecular data in our matrices (Fig.~\ref{Fig_Results-permeth_perparam} and Fig.~\ref{Fig_Results-global_perparam}). However, although these matrices are an order of magnitude smaller than some published matrices \citep[e.g.][]{springermacroevolutionary2012,nithe2013}, they are still within the size range of more modestly-sized empirical matrices \citep[e.g.][]{kellymolecular2014, sallam2011craniodental}. Therefore, our simulations reflect realistic parameters. Homoplasy, on the other hand, is expected to increase with an increase in the number of morphological characters \citep{}. % NC: Missing a citation here. Also is this correct - increase and increase? 
However, the use of probabilistic methods (i.e. Maximum Likelihood or Bayesian) and the M\textit{k} model \citep{lewisa2001} has been previously demonstrated to partially resolve this issue \citep{wrightbayesian2014}.

\subsection{Combined effect of missing data parameters}
As expected, when combining the missing data parameters, our ability to recover the "best" tree topology is affected in the same way as for the parameters individually: the Normalised Robinson-Foulds distance and the Normalised Triplets distance are higher when all the missing data parameters have few missing data (i.e. $M_{L}$ = 0\%, $M_{F}$ = 0\%, $M_{C}$ = 0\%) and lower when they have a lot of missing data (i.e. $M_{L}$ = 75\%, $M_{F}$ = 75\% and $M_{C}$ = 75\%; Fig.~\ref{Fig_Results-global_perparam}). However, it is important to notice that the effect of each parameter is not additive. The number of missing living taxa with morphological data ($M_{L}$) and the overall number of missing morphological characters ($M_{C}$), have a bigger effect than the amount of missing data for the fossil taxa ($M_{F}$), and when both $M_{L}$ and $M_{C}$ reach 50\% of missing data, the matrix does not contain enough phylogenetic information for the fossil taxa to be placed with confidence in the tree (Fig.~\ref{Fig_Results-paircomp_within}). This has important implications for building Total Evidence phylogenies (see below).

@@

Another practical implication of our results regards the tree inference methods. Because the Bayesian consensus trees consistently recovered topologies closer to the "best" tree topology than the Maximum Likelihood trees, we advise using Bayesian consensus trees to fix the topology where necessary for further steps in tree inference, for example in Tip-Dating \citep{BEASTmaster} % NC: Also cite ronquist paper?, and are we going with Tip-Dating? If yes, capitalise it throughout. 
Note that, although the Bayesian posterior tree distribution performed poorly in recovering the "best" tree topology, this is due to difficulties comparing distributions of trees (see discussion above). Thus, we do not suggest discarding the Bayesian posterior tree distribution once the topology has been fixed, particularly because these trees will be invaluable for phylogenetic comparative analyses.% \citep[e.g.][]{}. %Bayesian PCM cite % Nc: Could probably cite Jetz et al again?

@@

Missing data in Total Evidence matrices is not a problem for recovering the "best" tree topology as long as enough living and fossil taxa in the matrix have data for overlapping morphological characters. When missing data increases in any of our missing data parameters ($M_{L}$, $M_{F}$ or $M_{C}$), it reduces support for the placement of fossil taxa and increases the displacement of wildcard taxa. Therefore we advise filling as many gaps in Total Evidence matrices as possible. Because this is difficult, if not impossible, for fossil data, we recommend coding as many morphological characters, for as many living taxa, as possible. Additionally, the topology of the Bayesian consensus trees, regardless the amount of missing data, were always closer to the "best" tree topology than the Maximum Likelihood trees. Therefore, we advise using Bayesian consensus tree topologies if topology must be fixed in a Total Evidence analysis, for example when using the Tip-Dating method. The results of our analyses are encouraging and show that it is possible to combine both neontological and palaeontological data in the same phylogeny despite issues of missing data. Hopefully, using these approaches will greatly improve our understanding of macroevolutionary processes through both time and space.
% NC: Last sentence may be a bit of a stretch as we don't really talk about how to use it for macroevolutionary studies.

@@

\section{Acknowledgments}
Thanks to Gavin Thomas, Fr\'{e}d\'{e}ric Delsuc, Emmanuel Douzery, Trevor Hodkinson, Andrew Jackson, and April Wright for useful comments on our simulation protocol. Thanks to Paddy Doyle, Graziano D'Innocenzo and Sean McGrath for assistance with the computer cluster. Simulations used the Lonsdale cluster maintained by the Trinity Centre for High Performance Computing and funded through grants from Science Foundation Ireland. This work was funded by a European Commission CORDIS Seventh Framework Programme (FP7) Marie Curie CIG grant (proposal number: 321696).

@@

%---------------------------------------------
%
%       Supplementary, separated
%
%---------------------------------------------
% NC: I had some issues compiling this so I've commented it out.
%Ignore this for 'fast' pdf building
\section{Supplementary Material}

    %\input{SupplementaryMaterial/SupplementaryMaterial}
    %\label{SupplementaryMaterial}
